Model Search in Relational Domains

### 2016-10-20 contributions.tex
learning text

\item We present an abstract model search algorithm that strictly 
generalizes the Boolean Constraint Propagation (BCP) procedure to
compute a fixed point over a relational domain.  Our abstract 
conflict analysis algorithm present a non-trivial lifting of the 
Unique Implication Point (UIP) procedure to relational abstract 
domains in order to learn a generalized reason for conflicts.

>>>>

\item We present an abstract model search algorithm that strictly 
generalizes the Boolean Constraint Propagation (BCP) procedure to
compute a fixed point over a relational domain.

\item We propose several domain specific decision heuristics, that 
exploits the expressivity of the underlying abstract domain to make 
effective decisions.

\item We present a non-trivial lifting of the conflict analysis procedure 
to relational abstract domains in order to learn a generalized reason 
for conflicts.

### 2016-10-20 learning.tex
figure 4
learning text

\paragraph {\textbf{Abstract Conflict Graph}}
The first-UIP algorithm in SAT solver works on a data structure called 
{\em implication graph} that computes a cut that suffices to produce a 
conflict.  The trail in ACDCL implicitly encodes a graph structure called 
{\em abstract conflict graph}, which records the dependencies between 
deductions made during abstract model search phase.  Nodes in the graph 
represent elements of the trail $\mathcal{T}$.  Nodes can be either 
{\em decision} node or {\em deduction} node. Edges can be extracted from
the transformers in the reason trail.  Incoming arrows to node $n$ 
indicate that the predecessors of $n$ are sufficient to deduce $n$.  
A decision node has no incoming edges.  
For example, consider a formula 
$\varphi:= (x=5 \wedge y=x \wedge y=y+x \wedge y \leq 0)$ with current 
abstract valuation $v:= (x=5)$.  Figure~\ref{conflict} shows a 
snapshot of an abstract conflict graph that 
stores the deductions obtained from abstract model search using octagon domain.  
The last deduction conflict with the constraint $(y \leq 0)$.

\paragraph {\textbf{Lifting UIP to relational domains}}
An UIP is a special node in the abstract conflict graph such that 
any path from the last decision node to the conflict node must pass 
through it.  Unlike a SAT solver, a UIP in ACDCL may have meet irreducibles 
that involve the same variables.  The first-UIP is a node closest \pscmt{unique?} to the 
conflict, and the last UIP is the decision node itself.  Computing UIPs ensures 
asserting cuts, that is, it yields clauses that generates new deduction after 
backtracking.  Every cut in the graph reason for conflict that can be used in learning.  
A first-UIP algorithm traverses the trail $\mathcal{T}$ starting from the 
conflict node and iteratively finds new nodes that contradict \pscmt{we have to define what that means in the preliminaries} with the 
transformer stored in $\mathcal{R}[\bot]$.

For our example, there exists multiple incomaparable reasons for conflict,
marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
Choosing cut0 yields a learnt clause \pscmt{use $\leq$ instead of $<$} 
$L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$, 
which is obtained by negating the reason for conflict.  
%
\begin{figure}[t]
%\caption{An example of Octagon}\label{octagon}
\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
\caption{\label{conflict}
\pscmt{push constants to the lhs of the inequalities; labels cut0, cut1 not fully visible}
}
\end{figure} 
%    
\paragraph{\textbf{Learning in Relational Lattice}}
We model learning as a meet irreducible deduction transformer \pscmt{???}, called 
{\em abstract unit transformer (AUnit)} \pscmt{why the name? seems confusing as we are talking about learning and not about deduction}.  For an 
abstract lattice $A$ with complementable meet irreducibles 
and a set of conflicting meet irreducibles (countermodels) \pscmt{The term countermodels for $C$ is misleading since $\bigsqcap C$ is a set of countermodels, but the elements of $C$ individually are no countermodels.}
$C \subseteq A$ of $\varphi$ such that $\bigsqcap C$ does not 
satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is defined as follows. 
\pscmt{explain which cases are different in non-relational domains}
\[ AUnit(a) =
 \begin{cases}
  \bot       & \quad \text{if } a \sqsubseteq \meet{C} \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C
  \setminus \{t\}. a  \sqsubseteq t' \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C \setminus \{t\}. a
  \not\sqsubseteq t', \; \text{then} \\ 
             & \quad \forall a' \subseteq decomp(a) \;  \text{and} \; \forall a'' \in a \setminus \{a'\}, \\ 
             & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a' \not\sqsubseteq t  \\
  \top & \quad \text{otherwise} \\
 \end{cases}
\]
For example, let $C = \langle x+z \geq 10,  k \geq 0 \rangle$ \pscmt{why suddenly a different notation for $k$? The variable name $k$ might be confusing because it is usually used for counting or indices} be the conflicting set 
of meet irreducibles of $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
x+z \leq 15, k:[1,1] \rangle$.  Applying the third rule of $AUnit_C$, we get a 
decomposition of $a$ into $a'$ and $a''$ where 
$a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
$a''= \langle k:[1,1] \rangle \sqsubseteq (k \geq 0)$.  Then, $AUnit_C(a) = (x+z < 10)$, 
which constrains the bound of the octagonal constraint.  
Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$
and let $a = \langle x:[3,4], y:[5,6] \rangle$, then $AUnit_C(a) = \bot$ using
the first rule, since $a \sqsubseteq C$.  Now, let $a' = \langle x:[3,4] \rangle$, 
then $AUnit_C(a') = (y > 7)$ using second rule, since $a' \sqsubseteq (x \geq 2)$
and $a' \sqsubseteq (x \leq 5)$.

>>>>

\paragraph {\textbf{Abstract Conflict Graph}}
The first-UIP algorithm in SAT solver works on a data structure called 
{\em implication graph} that computes a cut that suffices to produce a 
conflict.  The trail in ACDCL implicitly encodes a graph structure called 
{\em abstract conflict graph}, which records the dependencies between 
deductions made during abstract model search phase.  Nodes in the graph 
represent elements of the trail $\mathcal{T}$.  Nodes can be either 
{\em decision} node or {\em deduction} node. Edges can be extracted from
the transformers in the reason trail.  Incoming arrows to node $n$ 
indicate that the predecessors of $n$ are sufficient to deduce $n$.  
A decision node has no incoming edges.  
For example, consider a formula 
$\varphi:= (x=5 \; \wedge \; y=x \; \wedge \; y=y+x \; \wedge \; y \leq 0)$ with current 
abstract valuation $v:= (x=5)$.  Figure~\ref{conflict} shows a 
snapshot of an abstract conflict graph that 
stores the deductions obtained from abstract model search using octagon domain.  
The last deduction conflict with the constraint $(y \leq 0)$.

\paragraph {\textbf{Lifting UIP to relational domains}}
An UIP is a special node in the abstract conflict graph such that 
any path from the last decision node to the conflict node must pass 
through it.  Unlike a SAT solver, a UIP for relational domain may 
involve meet irreducibles that contains the same variables.  The 
first-UIP is a unique node closest to the conflict, and the last UIP 
is the decision node itself.  Computing UIPs ensures asserting
cuts~\cite{cdcl,DBLP:journals/fmsd/BrainDGHK14}, that is, it 
yields clauses that generates new deduction after backtracking.  
Every cut in the graph is a reason for conflict that can be 
used in learning.  An abstract first-UIP algorithm~\cite{DBLP:journals/fmsd/BrainDGHK14} 
traverses the trail $\mathcal{T}$ starting from the conflict node and 
computes a cut that suffices to produce a conflict. 
\Omit {
contradict with the transformer stored in $\mathcal{R}[\bot]$.  
}
For our example, there exists multiple incomaparable reasons for conflict,
marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
Choosing cut0 yields a learnt clause 
$(x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$, 
which is obtained by negating the reason for conflict.  
%
\begin{figure}[t]
%\caption{An example of Octagon}\label{octagon}
\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
\caption{\label{conflict} Abstract Conflict Graph}
\end{figure} 
%    
\paragraph{\textbf{Learning in Relational Lattice}}
Learning in propositional solver yields a asserting clause~\cite{cdcl} 
that express the negation of the conflict reasons.  However, we model 
learning in relational domains as learning a transformer which infers 
new meet irreducibles using the abstract unit rule called 
{\em abstract unit transformer (AUnit)}.  We add $AUnit$ to the set of 
transformers \rmcmt{give symbol} used during model search. $AUnit$ is 
a generalization of propositional unit rule for numerical domains.  For 
an abstract lattice $A$ with complementable meet irreducibles and a set 
of meet irreducibles $C \subseteq A$ of $\varphi$ such that $\bigsqcap C$ does not 
satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is formally defined as follows. 
\[ AUnit(a) =
 \begin{cases}
  \bot       & \quad \text{if } a \sqsubseteq \meet{C} \qquad \qquad \qquad
  \qquad \qquad \qquad \quad (1)\\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C
  \setminus \{t\}. a  \sqsubseteq t' \qquad \quad (2) \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C \setminus \{t\}. a
  \not\sqsubseteq t', \; \text{then} \\ 
             & \quad \forall a' \subseteq decomp(a) \;  \text{and} \; \forall a'' \in a \setminus \{a'\}, \\ 
             & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a'
             \not\sqsubseteq t   \qquad \qquad \qquad \qquad (3) \\
  \top & \quad \text{otherwise} \qquad \qquad \qquad \qquad \qquad \qquad \quad (4) \\
 \end{cases}
\]

\pscmt{explain which cases are different in non-relational domains}
For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ 
be a conflicting element that does not satisfy $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
x+z \leq 15, w \geq 1, w \leq 1 \rangle$.  Applying the rule (3) of $AUnit_C$, we get a 
decomposition of $a$ into $a'$ and $a''$ where 
$a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
$a''= \langle w \geq 1, w \leq 1 \rangle \sqsubseteq (w \geq 0)$.  Then, $AUnit_C(a) = (x+z < 10)$, 
which constrains the bound of the octagonal constraint.  
Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$
and let $a = \langle x \geq 3, x \leq 4, y \geq 5, y \leq 6 \rangle$, then $AUnit_C(a) = \bot$ using
the first rule, since $a \sqsubseteq C$.  Now, let $a' = \langle x \geq 3, x \leq 4 \rangle$, 
then $AUnit_C(a') = (y > 7)$ using second rule, since $a' \sqsubseteq (x \geq 2)$
and $a' \sqsubseteq (x \leq 5)$.
